Revealing transient strain in geodetic data with Gaussian process regression

### 2017-07-18 RevisedSubmission/hines_hetland_2017_gji.tex
added more

\section{Discussion}\label{sec:Discussion}
Our results demonstrate that GPR is an effective tool for estimating transient strain from GNSS data, which can then be used to detect geophysical processes. Alternatively, one could detect geophysical signal by merely inspecting raw GNSS displacement time series. Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce visually discernible displacements in the GNSS time series. However, the GNSS data also contains outliers and non-tectonic deformation that is localized to individual stations. In contrast, our estimates of transient strain only identify features that are sufficiently spatially and temporally coherent, based on our chosen prior model. Furthermore, our estimates of transient strain are insensitive to common mode noise, which is highly spatially correlated noise resulting from factors such as reference frame error. Common mode noise can obscure geophysical signal in the GNSS data, but it gets canceled out when computing the spatial gradients to obtain transient strain. Lastly, if we want to detect transient geophysical signal for the purpose of assessing seismic hazard, then strain is perhaps a more meaningful metric to monitor than displacements.

>>>>

\section{Discussion}\label{sec:Discussion}
Our results demonstrate that GPR is an effective tool for estimating transient strain from GNSS data, which can then be used to detect geophysical processes. One may argue that geophysical signal can also be detected by merely inspecting the GNSS displacement time series. Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce visible displacements in the GNSS data. However, the GNSS data also contains outliers and non-tectonic deformation that is localized to individual stations. In contrast, our estimates of transient strain only identify features that are sufficiently spatially and temporally coherent, based on our chosen prior model. Furthermore, our estimates of transient strain are insensitive to common mode noise, which is highly spatially correlated noise resulting from factors such as reference frame error. Common mode noise can obscure geophysical signal in the GNSS data, but it gets canceled out when computing the transient strain. Lastly, our estimates of transient strain rates are a spatial and temporal derivative of displacements, and thus any geophysical signal in the transient strain rates tends to be more pronounced than in the GNSS data. For these reasons, we argue that transient strain rates estimated with the method described in Section \ref{sec:Method} can illuminate geophysical signal that may not be discernible from the noise in the GNSS displacement data.
